---
title: "Tidy Forecasting\\newline in R"
author: "Rob J Hyndman"
date: "ISF 2018"
fontsize: 14pt
output:
  beamer_presentation:
    fig_height: 4.33
    fig_width: 6
    highlight: tango
    keep_tex: no
    incremental: no
    theme: metropolis
    includes:
      in_header: preamble.tex
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  dev.args=list(bg=grey(0.9), pointsize=11)
)
savepdf <- function(file, width=16, height=16*4.33/6)
{
  fname <<- paste("figs/",file,".pdf",sep="")
  pdf(fname, width=width/2.54, height=height/2.54, pointsize=10)
  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
}
endpdf <- function()
{
  dev.off()
#  crop::dev.off.crop(fname)
}
set.seed(28052018)
options(digits=3,width=60)
```


# Why change?

## forecast package
\fontsize{13}{14}\sf

\hspace*{-0.6cm}\begin{tabular}{rl}
Pre 2003         & Private functions used for consulting projects \\
July/August 2003 & \textbf{ets} and \textbf{thetaf} added\\
%2003--2006      & Package available at robjhyndman.com \\
August 2006      & \textcolor{orange}{v1.0} available on CRAN \\
May 2007         & \textbf{auto.arima} added\\
%July 2008       & JSS paper (Hyndman \& Khandakar)\\
%September 2009  & \textcolor{orange}{v2.0}. Unbundled \\
May 2010         & \textbf{arfima} added\\
Feb/March 2011   & \textbf{tslm}, \textbf{stlf}, \textbf{naive}, \textbf{snaive} added \\
August 2011      & \textcolor{orange}{v3.0}. Box Cox transformations added\\
December 2011    & \textbf{tbats} added\\
April 2012       & Package moved to github\\
November 2012    & \textcolor{orange}{v4.0}. \textbf{nnetar} added\\
June 2013        & Major speed-up of \textbf{ets}\\
%December 2015   & 264,000 package downloads in one month!\\
February 2016    & \textcolor{orange}{v7.0}. Added ggplot2 graphics\\
February 2017    & \textcolor{orange}{v8.0}. Added \textbf{checkresiduals}, \textbf{tsCV} and \textbf{\%>\%}\\
April 2018       & \textcolor{orange}{v8.3}. Added \textbf{mstl}\\
June 2018        & $\approx$ 100,000 package downloads per month
\end{tabular}

## fable package
\placefig{10}{.1}{width=1.7cm}{fablesticker}

\begin{alertblock}{}
A replacement for the forecast package.
\end{alertblock}

### Why change?

 * Integrating with tidyverse packages
 * Designed for forecasting many related time series
 * Consistency of interface using formulas
 * Distribution forecasting rather than point + interval
 * Flexible transformations
 * Sub-daily data and multiple seasonal data handled more easily
 * Simpler interface for forecast reconciliation

# Example: Australian eating-out expenditure
## Example: Australian eating-out expenditure
\fontsize{10}{10}\sf

```{r auscafe1, echo=TRUE}
fpp2::auscafe
```

## Example: Australian eating-out expenditure
\fontsize{12}{12}\sf

```{r auscafe2, echo=TRUE}
library(tsibble)
cafe <- as_tsibble(fpp2::auscafe)
cafe
```

## Example: Australian eating-out expenditure
\fontsize{11}{11}\sf

```{r auscafe3, echo=TRUE, dependson='auscafe2'}
library(fable)
cafe %>% ETS(value)
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{10}{10}\sf

```{r auscafe4, echo=TRUE, dependson='auscafe2'}
cafe %>% ETS(value) %>% summary()
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{11}{11}\sf

```{r auscafe5, echo=TRUE, dependson='auscafe2'}
cafe %>% ETS(value) %>% forecast()
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{12}{12}\sf

```{r auscafe6, echo=TRUE, dependson='auscafe2'}
cafe %>% ETS(value) %>% forecast() %>%
  summary()
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{12}{12}\sf

```{r auscafe6a, echo=TRUE, dependson='auscafe2'}
cafe %>% ETS(value) %>% forecast() %>%
  summary(level=90)
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{11}{11}\sf

```{r auscafe7, echo=TRUE, dependson='auscafe2', fig.height=4}
cafe %>% ETS(value) %>% forecast() %>% autoplot()
```

\vspace*{10cm}

## Example: Australian eating-out expenditure
\fontsize{11}{11}\sf

```{r auscafe8, echo=TRUE, dependson='auscafe2'}
cafe %>% ARIMA(log(value)) %>%
  forecast() %>% summary()
```

# Example: Australian prison population

## Example: Australian prison population
\fontsize{11}{13}\sf

```{r prison1}
fpp2::prisonLF
```

## Example: Australian prison population
\fontsize{11}{13}\sf

```{r prison2}
prison <- fpp2::prisonLF %>%
  mutate(qtr=yearquarter(t)) %>%
  select(-t) %>%
  as_tsibble(index=qtr, key=id(state,gender,legal))
prison
```

## Example: Australian prison population
\fontsize{9}{11}\sf

```{r prison3, dependson='prison2'}
prison %>% ETS(count)
```

\vspace*{10cm}

## Example: Australian prison population
\fontsize{9}{11}\sf

```{r width1, include=FALSE}
options(width=151)
```

```{r prison4, dependson=c('prison2','width1')}
prison %>% ETS(count) %>% forecast()
```


```{r width2, include=FALSE}
options(width=60)
```

\fontsize{12}{13}\sf
###
Aggregation and reconciliation not yet implemented.

# Example: Half-hourly electricity demand
## Example: Half-hourly electricity demand
\fontsize{11}{14}\sf

```{r elecdemandsetup, echo=FALSE}
elecdemand <- as_tsibble(fpp2::elecdemand) %>%
  spread(key, value)
```

```{r elecdemand1, dependson="elecdemandsetup"}
elecdemand
```
## Example: Half-hourly electricity demand
\fontsize{10}{9.5}\sf

```{r elecdemand2, dependson="elecdemandsetup"}
fit2 <- ARIMA(elecdemand,
  Demand ~ Temperature + I(Temperature^2) + WorkDay)
summary(fit2)
```
```r
forecast(fit2, newdata=elecdemandfuture) %>% autoplot()
```

# Equivalent methods
## Equivalent methods: forecast $\longrightarrow$ fable

\hspace*{-0.7cm}\begin{tabular}{l@{~$\longrightarrow$~}p{8cm}}
\texttt{auto.arima} & \texttt{ARIMA} \\
\texttt{ets} & \texttt{ETS} \\
\texttt{tslm/lm} & \texttt{LM} \\
\texttt{tbats} & \texttt{TBATS} \\
\texttt{nnetar} & \texttt{NNAR} \\
\texttt{stlm} & \texttt{STL \%>\%} \par
 \hspace*{0.2cm}\texttt{modelcomponents(} \par
 \hspace*{0.6cm}\texttt{ETS(seasadj),SNAIVE(season))}
\end{tabular}


###
 * All functions have a formula interface with automatic modelling if no formula provided.
 * All functions produce `mable` class objects.
 * Some of these functions not yet implemented

## Equivalent methods: forecast $\longrightarrow$ fable
\fontsize{14}{15}\sf

\hspace*{-0.7cm}\begin{tabular}{l@{~$\longrightarrow$~}p{10cm}}
\texttt{naive} &  \texttt{NAIVE \%>\% forecast} \\
\texttt{snaive} &  \texttt{SNAIVE \%>\% forecast} \\
\texttt{thetaf} &  \texttt{THETA \%>\% forecast} \\
\texttt{stlf} &  \texttt{STL \%>\%} \par
 \hspace*{0.3cm}\texttt{modelcomponents(} \par
 \hspace*{0.6cm}\texttt{ETS(seasadj),SNAIVE(season)) \%>\%} \par
 \hspace*{0.3cm}\texttt{forecast} \\
\texttt{hw} &  \texttt{HW \%>\% forecast} \\
\texttt{holt} & \texttt{HOLT \%>\% forecast} \\
\texttt{ses} &  \texttt{SES \%>\% forecast} \\
\texttt{splinef} & \texttt{SPLINE \%>\% forecast} \\
\texttt{croston} & \texttt{CROSTON \%>\% forecast}
\end{tabular}

###
 * `forecast` produces `fable` class objects.

# Extensibility

## Extending fable

### Tools to easily create new fable models
 * Easily create specials for model formulae
 * Focus on model estimation and forecasts
 
### Automatically supported fable functionality
 * Transformations and back-transformations (with bias adjustments)
 * Prediction intervals
 * Plotting tools
 * Accuracy measures and evaluation
 * Model combinations (hierarchies & ensembles)

# More information

## More information
\fontsize{12.8}{15}\sf
\placefig{8}{.1}{width=1.7cm}{tsibblesticker}
\placefig{10}{.1}{width=1.7cm}{fablesticker}
\vspace*{.5cm}

###
```r
devtools::install_github("tidyverts/tsibble")
devtools::install_github("tidyverts/fable")
```

###
\fontsize{11}{11}\sf
\centering\begin{tabular}{p{3.35cm}p{3.cm}p{3.5cm}}
\includegraphics[height=3.25cm]{dicook} &
\includegraphics[height=3cm]{earowang} &
\includegraphics[height=3cm]{mitch}\\
Di Cook & Earo Wang  & Mitchell O'Hara-Wild
\end{tabular}


### Follow our progress
  * `tidyverts.org`
  * `robjhyndman.com/hyndsight`
